---
title: "Analysis of SARS-CoV-2 SNP Data"
author: "Kim Medina"
date: "2024-09-23"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

# Load Packages and Data

We are going to load both the dplyr and ggplot2 packages for our analysis.

```{r}
library(dplyr)
library(ggplot2)
library(knitr)

# load in SNP data
snp_data <- read.csv("data/sample_snp_data.csv")
```

The first question we have about this dataset is the quality of the SNP calls for each of the three different samples in this dataset. To assess this we'll take a look at the mean and standard deviation of the quality scores in the dataset.

```{r, summarize-quality}
snp_data %>%
  group_by(sample) %>%
  summarise(mean_quality = mean(qual),
            sd_quality = sd(qual)) %>%
  kable(digits = 2)
```

From this, we can see that sample SRR12433065 has a lower mean quality and higher standard deviation that the other two samples. This means it might be a little less robust and perhaps we may want to consider redoing it or excluding it.

To dig in a bit here, let's make a plot looking at the distribution of SNPs across the genome for these three sample to see if there is some pattern. Maybe all the low quality ones are in a specific region.

```{r, plot-quality-across-genome}
snp_data %>%
  ggplot(aes(x = pos,
             y = qual,
             color = sample)) + 
  geom_point(size = 3) +
  labs(title = "SNP quality across SARS-CoV-2 Genome",
       x = "Position in Genome (nt)",
       y = "SNP Quality")
```

It looks like from the plot, the lower quality SNPs are distributed throughout the genome, so it is not just limited to one area. There are also a few high quality SNPs in the otherwise generally low-quality sample (SRR12433065) at the beginning of the genome.
